Automated troubleshooting toolset for SDNs

### 2013-10-27 draft.tex
NICE

\section{SDN layering, the key used to a better troubleshooting}\label{layering}
 % Expliquer ici la stack SDN (decomposition en layer) et donner l'intuition concernant la plus grande facilité pour remonter à l'origine d'un probleme.
Finding and solving network bugs are not the aim of SDN, but we can use it to re-think the way we troubleshoot networks.\\ The SDN architecture is decomposed into  layers, those layers can be represented in a two dimensionnal array. As you can see on Figure \ref{layers}, we have the two main layers called \textit{State layers} and \textit{Code layers}. The state layers hold a representation of the network's configuration for each parts of the network architecture. The code layers implement logic to maintain the mapping between two state layers. Each states layers should verify the equivalence properties, which means that each of them should correctly mapping every other state layer. The idea is that, for each policy, if the state layers are correctly mapped among each other, then the policy is  set and acts like it should.\\
On the Figure \ref{layers}, you have the following elements. \\
\textbf{Policy} - Policies are set up by the network administrator to configure the Logical View. These policies can be routing, access control or QoS policies. They are written inside a control application. See \cite{...} for exemple.
\\
\textbf{Logical View} - Abstract representation of the network which aims to make things easier for the control application to create policies. A mapping between the Logical View and one or more Physical View are done by Network Hypervisor.\\
\textbf{Physical View} - This is a correspondance with the real network element, a representation of it handled by the network OS. The protocol used to configure the network element is one like OpenFlow \cite{...https://www.opennetworking.org/}. A physical View has thus a one-to-one mapping with a real network element.
\\
\textbf{Device State} - State of the network element maintened by its firmware.
\\
\textbf{Hardware} - Network element. \\

@@

\subsection{VeriFlow}
If the actual behavior of the network does not correspond to the policy, then the issue can potentially be found in the erroneous correspondance between the \textit{device state layer} and the \textit{policy layer}. There exist different tools to verify this hypothesis. The tools we discuss here have the advantage to verify some network invariants in real-time. Other tools as Anteater and Header Space Analysis (HSA) that are also used in the same purpose carry out the static analysis of snapshots of the network data-plane state\cite{Kazemian:2012:HSA:2228298.2228311}\cite{Mai:2011:DDP:2018436.2018470}. Since SDN controllers are capable of installing around 30.000 flows per second while maintaining less than 10 ms delay for the installation, it is absolutely not enough to have tools that check for network invariant with a latency of the order of seconds \cite{Khurshid:2013:VVN:2482626.2482630}.\\
VeriFlow acts as a proxy in between the SDN controller and the switches that verifies some common network invariants at each forwarding rule installation and that at very high speed. In order to achieve a 7ms delay inflation with regards to traditional TCP connections\footnote{Test configuration : 20 node OpenFlow network sing Mininet and a NOX controller\cite{Khurshid:2013:VVN:2482626.2482630} }, VeriFlow slices the network into equivalences classes, then builds a virtual forwarding graph for each ones using a trie structure, and finally checks invariants by traversing those graph with a depth-first search approach. Note that VeriFlow allows to spotting network issues before they reach the network, which is very valuable.\\
As limitations, we can note that presently, VeriFlow only checks for reachability invariants, and that it is not suitable for multiple/distributed controller \cite{Khurshid:2013:VVN:2482626.2482630}.

@@

>>>>

\section{SDN layering, the key used to an automated troubleshooting}\label{layering}
 % Expliquer ici la stack SDN (decomposition en layer) et donner l'intuition concernant la plus grande facilité pour remonter à l'origine d'un probleme.
Finding and solving network bugs are not the aim of SDN, but we can use it to re-think the way we troubleshoot networks.\\ The SDN architecture is decomposed into  layers, those layers can be represented in a two dimensionnal array. As you can see on Figure \ref{layers}, we have the two main layers called \textit{State layers} and \textit{Code layers}. The state layers hold a representation of the network's configuration for each parts of the network architecture. The code layers implement logic to maintain the mapping between two state layers. Each states layers should verify the equivalence properties, which means that each of them should correctly mapping every other state layer. The idea is that, for each policy, if the state layers are correctly mapped among each other, then the policy is  set and acts like it should.\\
On the Figure \ref{layers}, you have the following elements. \\
\textbf{Policy} - Policies are set up by the network administrator to configure the Logical View. These policies can be routing, access control or QoS policies. They are written inside a control application. See \cite{...} for exemple.
\\
\textbf{Logical View} - Abstract representation of the network which aims to make things easier for the control application to create policies. A mapping between the Logical View and one or more Physical View are done by Network Hypervisor.\\
\textbf{Physical View} - This is a correspondance with the real network element, a representation of it handled by the network OS. The protocol used to configure the network element is one like OpenFlow \cite{...https://www.opennetworking.org/}. A physical View has thus a one-to-one mapping with a real network element.
\\
\textbf{Device State} - State of the network element maintened by its firmware.
\\
\textbf{Hardware} - Network element. \\

@@

\subsection{VeriFlow}
If the actual behavior of the network does not correspond to the policy, then the issue can potentially be found in the erroneous correspondance between the \textit{device state layer} and the \textit{policy layer}. There exist different tools to verify this hypothesis. The one we discuss here have the advantage to verify some network invariants in real-time. Other tools as Anteater and Header Space Analysis (HSA) that are also used in the same purpose carry out the static analysis of snapshots of the network data-plane state\cite{Kazemian:2012:HSA:2228298.2228311}\cite{Mai:2011:DDP:2018436.2018470}. Since SDN controllers are capable of installing around 30.000 flows per second while maintaining less than 10 ms delay for the installation, it is absolutely not enough to have tools that check for network invariant with a latency of the order of seconds \cite{Khurshid:2013:VVN:2482626.2482630}.\\
VeriFlow acts as a proxy in between the SDN controller and the switches and this proxy verifies some common network invariants at each forwarding rule installation with a very high speed. In order to achieve a 7ms delay inflation with regards to traditional TCP connections\footnote{Test configuration : 20 nodes OpenFlow network sing Mininet and a NOX controller\cite{Khurshid:2013:VVN:2482626.2482630} }, VeriFlow slices the network into equivalences classes, then builds a virtual forwarding graph for each ones using a trie structure, and finally checks invariants by traversing those graph with a depth-first search approach. Note that VeriFlow allows to spotting network issues before they reach the network, which is very valuable.\\
As limitations, we can note that presently, VeriFlow only checks for reachability invariants, and that it is not suitable for multiple/distributed controller \cite{Khurshid:2013:VVN:2482626.2482630}.

@@

Following the SDN layering technique, once the erroneous 2-layers correspondance has been determined, another tool can be used to spot the bug more precisely, either in the controller software or in the firmware running on the switches.\\
The NICE tool is used to produce traces leading to a bug, by systematically exploring the set of possible states of the system (including controller, switches and hosts) and checking them regards to some network invariants\cite{Canini:2012:NWT:2228298.2228312}. What makes this contribution to OpenFlow verification a real asset is its ability to deal with a unbound space state. The real challenge for testing openFlow applications is the scalability.
